Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-08-08 PAPER.tex
edited signal loss methodology

One promising technique to probe the EoR is to target the 21 cm wavelength emission that is emitted by neutral hydrogen via its spin-flip transition. This technique is powerful because it can be observed as a function of redshift --- that is, the wavelength of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before stretching out as it traveled through expanding space. The 21 cm line offers a window into the evolution of ionization, temperature, and density fluctuations on cosmic scales.

@@

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and instrumental systematics that are 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, traps, and connections to the themes . We first approach the themes from a broad perspective, and then perform a detailed case study using data from the 64-element configuration of PAPER.

This paper is organized as follows. In section 2 we introduce the three themes of our focus, using a toy model to develop intuition into each one. Section 3 presents a case study into each theme using data from the PAPER-64 array, highlighting key changes from \citet{ali_et_al2015}. We conclude in Section 7.

@@

There are many choices for a data analyst. How can time-ordered measurements be combined? How can the variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not destroying an EoR signal? How can the source of a detection be properly identified? Many common techniques, such as averaging data, weighting, bootstrapping, and jackknife testing, each address these issues yet harbor additional trade-offs. For example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

Despite there being many data analysis choices, measuring the statistical $21$ cm power spectrum ultimately requires robust methods for determining accurate confidence intervals and rigorous techniques to identify and suppress systematics.  In this paper, we focus on three $21$ cm power spectrum themes that encapsulate this goal and discuss four techniques that interplay with each other and the themes. We will give brief definitions now, and build intuition for each theme in the sections to follow.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power spectrum result. Stemmed from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): As explained in the next section, certain analysis techniques can lead to the loss of the EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss has subtle challenges but is crucial to the accuracy of any result.
\item \textbf{Error Bar Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result determine the difference between a detection and a null result, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of bias in a visibility measurement that can leak its way into a power spectrum, such as bias from noise and foregrounds. Proving a bias is EoR may be the most difficult challenge for $21$ cm analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper we will highlight some sources of bias and discuss ways to mitigate their effects.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. We focus on these techniques because of their strong appeal for a power spectrum analysis. However, they also have limitations, and we will discuss circumstances in which there are trade-offs.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering can be considered an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which weights a dataset by minimizing the covariance between frequency channels. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, one challenge of weighting is in accurately determining a covariance matrix that best describes the data.
\item \textbf{Bootstrapping:} Because applying theoretical errors to a $21$ cm measurement are insufficient \cc{why?}, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we get a sense of its inherent variance, though there are subtleties to consider such as bootstrapping sample sizes.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}).
\end{itemize}

@@

Driven by the need to mitigate foreground bias, we would like to use a weighting method that succeeds in down-weighting foregrounds. We apply this weighting to data and propagate it into a final estimator using the power spectrum estimation technique of optimal quadratic estimators (OQE; \citealt{liu_tegmark2011}; \citealt{liu_et_al2014b}). Before showing how signal loss can arise when using different weighting matrices, we first summarize OQE as follows.

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities in Jy that have the shape ($times$, $frequencies$). We form the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{S}\textbf{Q}_{\alpha}\textbf{S}\textbf{x},
\end{equation}

\noindent where $\textbf{S}$ is a weighting matrix (we use $\textbf{S}$ since $\textbf{W}$ is reserved for window functions). As an example, inverse covariance weighting would set $\textbf{S} = \textbf{C}^{-1}$ and an unweighted case would use $\textbf{S} = \textbf{I}$, the identity matrix. The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into power spectrum space. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the Fourier-dual to frequency.

@@

\begin{equation}
\hat{\textbf{p}} = \textbf{M}\hat{\textbf{q}},
\end{equation}

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. We compute $\textbf{M}$ using the Fisher matrix $\textbf{F}$, defined as:

\begin{eqnarray}
\textbf{F}_{\alpha\beta} &=& \frac{1}{2}tr[\textbf{S}\textbf{Q}_{\alpha}\textbf{S}\textbf{Q}_{\beta}]. 
\end{eqnarray}

We have a choice for $\textbf{M}$. For simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis of PAPER-64 data as explained in Section \ref{sec:PS}

In the next sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{S}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds perfectly without destroying an EoR signal.

@@

One choice for a weight matrix $\textbf{S}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds. \cc{math!} The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. \cc{math!} However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting the foregrounds in the data.

@@

\begin{equation}
\hat{\textbf{C}} \equiv \langle\textbf{xx}^{\dagger}\rangle,
\end{equation}

\noindent assuming $\langle\textbf{x}\rangle = 0$, where $\langle \rangle$ denotes a time-average \cc{distinguish between ideal ensemble average and finite estimate???}. The inverse covariance matrix is therefore $\textbf{S} = \textbf{C}^{-1}$. \cc{should I put hat's on all C's that are estimated from the data? Need to distinguish it from ideal C.}

First, we compute the power spectrum of $\textbf{x}$ using the OQE formalism and a weighting matrix of $\textbf{C}^{-1}$. The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the unweighted power spectrum of $\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0.2cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymode} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data (left), contrasted with projecting out the first eigenmode only (right). There is signal loss for the former but not the latter. \cc{elaborate}}
	\label{fig:toy_sigloss3}
\end{figure*}

As shown, our inverse covariance-weighted result successfully suppresses foregrounds \cc{explain why}. This illustrates why inverse covariance weighting is attractive for $21$ cm data. It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In order to understand how this occurs, we need to closely study our covariance matrix, $\textbf{C}$, which is shown in Figure \ref{fig:toy_sigloss12}.

@@

If $\textbf{C}$ is computed from the data itself and its inverse is then applied to the data, it carries the risk of over-fitting information in the data and introducing a multiplicative bias to estimates of the signal \cc{elaborate with math!}. As a contradictory example, if instead we choose $\textbf{S} = \textbf{I}$, there will be no signal loss (although it may suffer from foreground or instrumental bias).

It turns out that because we estimated $\textbf{C}$ from our data, its eigenspectrum offers insight into understanding how inverse covariance weighting affects our result. An eigenspectrum ranks the eigenvalues of $\textbf{C}$ from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. Each eigenvalue encodes the amount of variance in the dataset for a particular mode. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. \cc{use more rigorous math to supplement this paragraph}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\textbf{C}_{FG}$ (blue), $\textbf{C}_{EoR}$ (red), and $\textbf{C}_{FG+EoR}$ (green). The eigenspectrum of $\textbf{C}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. All eigenspectra steepen slightly for the final few eigenmodes, a behavior that results in some modes being upweighted more strongly than others, thereby overfitting noise and causing signal loss in the final power spectrum estimate (Figure \ref{fig:toy_sigloss3}, left). If only the first eigenmode is used, all others get equal weights and hence, there is no signal loss (Figure \ref{fig:toy_sigloss3}, right). \cc{also illustrate first few and last few eigenvalues -- how to best show this?}}
	\label{fig:toy_sigloss2}
\end{figure}

For our toy model, the eigenspectra of $\textbf{C}_{FG}$, $\textbf{C}_{EoR}$, and $\textbf{C}$ are shown in Figure \ref{fig:toy_sigloss2}. The blue curve peaks only at the zeroth eigenmode, revealing that the foregrounds can be described by a single eigenmode. This is a direct consequence of how it was constructed (only one sinusoid). The eigenspectrum of EoR, however, is fairly flat, signifying that its covariance matrix is similar to an identity matrix. This makes sense because EoR consists of a random signal, which should only have strong frequency-frequency correlations when two frequencies are the same.

@@

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Altering $\textbf{C}$ in this way is one example of a regularization method, in which we are changing $\textbf{C}$ in a way that flattens its eigenspectrum, making it more identical to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds, we can down-weight them without signal loss. There are several other ways to regularize $\textbf{C}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

In Section \ref{sec:toymodel} we showed how altering $\textbf{C}$ can make the difference between zero and some signal loss, if we can distinguish between foreground eigenmodes and EoR eigenmodes. We will now use our toy model to describe several other ways to tailor $\textbf{C}$ in order to minimize signal loss. We illustrate the resulting power spectra and eigenspectra for four different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}.

As a first test, we know that our simulated EoR should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such, instead of computing it based on the dataset itself, and add it to $\textbf{C}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$ to obtain a final $\textbf{C}$ to use in weighting, we see that there is negligible signal loss (Figure \ref{fig:toy_sigloss8}, upper left). This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\textbf{C}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves differ at one $k$ value in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices.

\begin{figure*}
	\centering
	\includegraphics[trim={0.4cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={1cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0.4cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={1cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each avoid signal loss, including modeling the covariance matrix of EoR (upper left), regularizing $\textbf{C}$ by adding an identity matrix to it (upper right), using only the first three eigenmodes of $\textbf{C}$ (lower left), and multiplying an identity matrix to $\textbf{C}$ (lower right). There are many regularization options available to minimize signal loss, and some work better than others depending on the dataset.}
	\label{fig:toy_sigloss8}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\textbf{C}$ (green) to that of four alternate weighting options, including modeling the covariance matrix of EoR (red), regularizing $\textbf{C}$ by adding an identity matrix to it (blue), using only the first three eigenmodes of $\textbf{C}$ (yellow), and multiplying an identity matrix to $\textbf{C}$ (magenta). }
	\label{fig:toy_sigloss14}
\end{figure}

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\textbf{C} = \textbf{C} + \textbf{I}$. By adding the identity, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum \cc{why?}.

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} flattens out the eigenspectrum of $\textbf{C}$ a different way - by zeroing all but the first three eigenmodes. This means that the first three eigenmodes will be given weights (the three highest variance modes will be down-weighted), but the rest will be treated with similar weights. Hence, we avoid up-weighting the last few noisy modes and have negligible signal loss. However, we do not perfectly recover the shape of EoR because we lost information when projecting out certain modes.

The last regularization scheme we are highlighting here is setting $\textbf{C} = \textbf{C} \cdot \textbf{I}$. In the bottom right panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by the identity results in a diagonal matrix, meaning we are only left with correlation information between the same two frequencies. Because we disregard information from all other frequency combinations, we do a poor job suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.

>>>>

One promising technique to probe the EoR is to target the 21 cm wavelength emission that is emitted by neutral hydrogen via its spin-flip transition. This technique is powerful because it can be observed as a function of redshift --- that is, the wavelength of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before stretching out as it traveled through expanding space. Tracing the 21 cm line as a function of redshift offers a window into the evolution of ionization, temperature, and density fluctuations during this transformative era.

@@

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and instrumental systematics that are, when combined, 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, traps, and connections to the themes . We first approach the themes from a broad perspective, and then perform a detailed case study using data from the 64-element configuration of PAPER.

This paper is organized as follows. In section 2 we introduce the three themes of our focus, using a toy model to develop intuition for each one. Section 3 presents a case study into each theme using data from the PAPER-64 array, highlighting key changes from \citet{ali_et_al2015}. We conclude in Section 7.

@@

There are many choices for a data analyst. How can time-ordered measurements be combined? How can the variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not destroying an EoR signal? How can the source of a detection be properly identified? Many common techniques, such as averaging data, weighting, bootstrapping, and jackknife testing, address these issues but harbor additional trade-offs. For example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

Despite there being many data analysis choices, measuring the statistical $21$ cm power spectrum ultimately requires robust methods for determining accurate confidence intervals and rigorous techniques to identify and suppress systematics.  In this paper, we focus on three $21$ cm power spectrum themes that encapsulate this goal and discuss four techniques that interplay with each other and impact the themes. We will give brief definitions now, and build intuition for each theme in the sections to follow.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power spectrum result. Stemmed from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): As explained in the next section, certain analysis techniques can lead to the loss of the EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss has subtle challenges but is crucial to the accuracy of any result.
\item \textbf{Error Bar Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result determine the difference between a detection and a null result, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of bias in a visibility measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. A successful EoR detection would also imitate a bias. Proving a bias is EoR may be the most difficult challenge for $21$ cm analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper we will highlight some sources of bias and discuss ways to mitigate their effects.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. We focus on these techniques because \cc{??}. However, some have limitations, and we will discuss circumstances in which there are trade-offs.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is considered an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which is a generalized version of inverse variance weighting that also takes into account data correlations. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance weighting is in accurately describing a covariance matrix that best describes the data. 
\item \textbf{Bootstrapping:} Because theoretical models for covariance matrices are questionable and theoretical error estimation methods rely on assumptions that can be violated by data, applying theoretical errors to a $21$ cm measurement is often avoided. Hence, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we get a sense of its inherent variance, though there are subtleties to consider such as the independence of values in a dataset.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}).
\end{itemize}

@@

Driven by the need to mitigate foreground bias, we use a weighting method that succeeds in down-weighting foregrounds. We apply this weighting to data and propagate it into a final estimator using the power spectrum estimation technique of optimal quadratic estimators (OQE; \citealt{liu_tegmark2011}; \citealt{liu_et_al2014b}). Before showing how signal loss can arise when using different weighting matrices, we first summarize OQE as follows.

We begin with our data vector, $\textbf{x}$ which contains our measured visibilities in Jy. It has length ($N_{t}, N_{f}$), where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of the Fourier transform of the sky along spatial dimensions, and since we are interested in $3$-dimensional Fourier-modes we only need to take one Fourier transform along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}_{\alpha}\textbf{R}\textbf{x},
\end{equation}

\noindent where $\textbf{R}$ is a weighting matrix. As an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $\textbf{R} = \textbf{I}$, the identity matrix. The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into power spectrum space while also taking into account physical constants such as the Boltzmann constant used to convert Jy to Kelvin. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}).

@@

\begin{equation}
\label{eq:phat}
\hat{\textbf{p}} = \textbf{M}\hat{\textbf{q}},
\end{equation}

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. We have a choice for $\textbf{M}$. For simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis of PAPER-64 data as explained in Section \ref{sec:Bias}

In the next sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds perfectly without destroying an EoR signal.

@@

One choice for a weight matrix $\textbf{R}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds and yields the smallest possible error bars on the measurements (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting the foregrounds in the data.

@@

\begin{equation}
\hat{\textbf{C}} \equiv \langle\textbf{xx}^{\dagger}\rangle_{t},
\end{equation}

\noindent assuming $\langle\textbf{x}\rangle = 0$, where $\langle \rangle$ denotes a finite time-average. The inverse covariance matrix is therefore $\textbf{R} = \hat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of $\textbf{x}$ using the OQE formalism and a weighting matrix of $\hat{\textbf{C}}^{-1}$. The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the unweighted power spectrum of $\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0.2cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data (left), contrasted with projecting out the first eigenmode only (right). There is signal loss for the former but not the latter. \cc{elaborate}}
	\label{fig:toy_sigloss3}
\end{figure*}

As shown, our inverse covariance-weighted result successfully suppresses foregrounds \cc{explain why}. It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In order to understand how this occurs, we need to closely study our covariance matrix, $\hat{\textbf{C}}$, which is shown in Figure \ref{fig:toy_sigloss12}.

@@

If $\textbf{C}$ is computed from the data itself and its inverse is then applied to the data, it carries the risk of over-fitting information in the data and introducing a multiplicative bias to estimates of the signal \cc{elaborate with math!}. As a contrasting example, if instead we choose $\textbf{R} = \textbf{I}$, there will be no signal loss (although it may suffer from foreground or instrumental bias).

It turns out that because we estimated $\textbf{C}$ from our data, its eigenspectrum offers insight into understanding how inverse covariance weighting affects our result. An eigenspectrum ranks the eigenvalues of $\hat{\textbf{C}}$ from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. Each eigenvalue encodes the amount of variance in the dataset for a particular mode. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. All eigenspectra steepen slightly for the final few eigenmodes, a behavior that results in some modes being upweighted more strongly than others, thereby overfitting noise and causing signal loss in the final power spectrum estimate (Figure \ref{fig:toy_sigloss3}, left). If only the first eigenmode is used, all others get equal weights and hence, there is no signal loss (Figure \ref{fig:toy_sigloss3}, right). \cc{also illustrate first few and last few eigenvalues -- how to best show this?}}
	\label{fig:toy_sigloss2}
\end{figure}

For our toy model, the eigenspectra of $\hat{\textbf{C}}_{FG}$, $\hat{\textbf{C}}_{EoR}$, and $\hat{\textbf{C}}$ are shown in Figure \ref{fig:toy_sigloss2}. The blue curve peaks only at the zeroth eigenmode, revealing that the foregrounds can be described by a single eigenmode. This is a direct consequence of how it was constructed (only one sinusoid). The eigenspectrum of EoR, however, is fairly flat, signifying that its covariance matrix is similar to an identity matrix. This makes sense because EoR consists of a random signal, which should only have strong frequency-frequency correlations when two frequencies are the same.

@@

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Altering $\hat{\textbf{C}}$ in this way is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more identical to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds, we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

In Section \ref{sec:toymodel} we showed how altering $\hat{\textbf{C}}$ can make the difference between zero and some signal loss, if we can distinguish between foreground eigenmodes and EoR eigenmodes. We will now use our toy model to describe several other ways to tailor $\hat{\textbf{C}}$ in order to minimize signal loss. We illustrate the resulting power spectra and eigenspectra for four different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}.

As a first test, we know that our simulated EoR should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such, instead of computing it based on the dataset itself, and add it to $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$ to obtain a final $\hat{\textbf{C}}$ to use in weighting, we see that there is negligible signal loss (Figure \ref{fig:toy_sigloss8}, upper left). This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves differ at one $k$ value in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices.

\begin{figure*}
	\centering
	\includegraphics[trim={0.4cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={1cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0.4cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={1cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each avoid signal loss, including modeling the covariance matrix of EoR (upper left), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (upper right), using only the first three eigenmodes of $\hat{\textbf{C}}$ (lower left), and multiplying an identity matrix to $\hat{\textbf{C}}$ (lower right). There are many regularization options available to minimize signal loss, and some work better than others depending on the dataset.}
	\label{fig:toy_sigloss8}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix to $\textbf{C}$ (magenta). }
	\label{fig:toy_sigloss14}
\end{figure}

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}} + \textbf{I}$. By adding the identity, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum \cc{why?}.

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} flattens out the eigenspectrum of $\hat{\textbf{C}}$ a different way - by zeroing all but the first three eigenmodes. This means that the first three eigenmodes will be given weights (the three highest variance modes will be down-weighted), but the rest will be treated with similar weights. Hence, we avoid up-weighting the last few noisy modes and have negligible signal loss. However, we do not perfectly recover the shape of EoR because we lost information when projecting out certain modes.

The last regularization scheme we are highlighting here is setting $\hat{\textbf{C}} = \hat{\textbf{C}} \cdot \textbf{I}$. In the bottom right panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by the identity results in a diagonal matrix, meaning we are only left with correlation information between the same two frequencies. Because we disregard information from all other frequency combinations, we do a poor job suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.
